### 2024-12-09 Draft1.tex
Added changes to the explanations of the different methods and the practical case

\setlength{\parindent}{0.5in} % Sets the paragraph indentation to 0.25 inches

@@

% Actual exploration
\section{Exploration}

\subsection*{Theoretical Foundation of Calculus Cryptography}

At its core, calculus cryptography utilizes differential equations as encryption mechanisms. A simple example involves encoding information as the solution to a specific differential equation. The sender encodes a message \( M \) by constructing a differential equation where \( M \) is embedded within the boundary or initial conditions. For example:

@@

Here, \( M \) serves as the initial condition. The encrypted data is transmitted in the form of \( f(x, y) \), and only someone with knowledge of \( M \) can correctly solve the equation to recover the message.

\subsubsection*{Key Components}
\begin{enumerate}
    \item \textbf{Encryption:} Represent the message \( M \) as an initial value in a carefully constructed differential equation.
    \item \textbf{Transmission:} Share the function \( f(x, y) \) and any necessary boundary constraints, keeping \( M \) private.
    \item \textbf{Decryption:} Solve the differential equation using the initial condition.
\end{enumerate}

\subsubsection*{Example}

Let \( f(x, y) = y + x^2 \), and the initial condition \( y(0) = 5 \). The differential equation is:

@@

\subsection*{Comparison with the Discrete Logarithmic Problem (DLP)}

The DLP is a fundamental problem in classical cryptography. It involves finding an integer \( k \) such that:

\[
g^k \equiv h \pmod{p}
\]

where \( g \) is a generator, \( h \) is the resultant, and \( p \) is a prime modulus. DLP-based systems, such as Diffie-Hellman and ElGamal, derive their security from the computational difficulty of solving this congruence.

\subsubsection*{Steps in the DLP}
\begin{enumerate}
    \item \textbf{Parameter Generation:} Select \( g \) and \( p \) such that \( g \) generates a large cyclic group under modulo \( p \).
    \item \textbf{Public Sharing:} Share \( g \), \( p \), and \( g^k \) while keeping \( k \) private.
    \item \textbf{Encryption and Decryption:} Use \( k \) as the secret key for modular exponentiation operations.
\end{enumerate}

\subsection*{Comparative Analysis: Mathematical Complexity and Security}

\subsubsection*{Mathematical Complexity}
\begin{itemize}
    \item \textbf{Calculus Cryptography:} Involves solving differential equations, typically requiring integration techniques such as substitution or partial fractions. These methods demand a solid grasp of calculus and provide an opportunity to explore more advanced topics like Laplace transforms or numerical methods.
    \item \textbf{DLP:} Requires modular arithmetic and group theory concepts. While computationally challenging due to its NP-hard nature, the mathematical foundation (exponentiation and congruences) is more elementary compared to solving advanced calculus problems.
\end{itemize}

\subsubsection*{Efficiency}
\begin{itemize}
    \item \textbf{Calculus Cryptography:} Less efficient due to the computational intensity of solving differential equations. For large-scale data, the reliance on numerical solutions further impacts performance.
    \item \textbf{DLP:} Highly efficient for encryption, with established algorithms like modular exponentiation. However, decryption and brute-forcing \( k \) remain computationally expensive.
\end{itemize}

\subsection*{Security}
\begin{itemize}
    \item \textbf{Calculus Cryptography:} Relies on the difficulty of reverse-engineering specific differential equations. While promising, it lacks the extensive cryptanalysis that validates the DLP's security.
    \item \textbf{DLP:} Proven resilience against traditional attacks like brute force or discrete logarithm computation within feasible bounds, making it a benchmark for cryptographic security.
\end{itemize}

@@

% Exploration of DLP
\section*{Exploration of a Practical Case: Discrete Logarithmic Problem (DLP)}

@@

\subsection*{Encryption}

@@

\section*{Exploration of a Practical Case: Calculus Cryptography}

To encrypt a large numerical message \( M \), calculus cryptography uses differential equations to encode the data into a function, with \( M \) embedded in its initial or boundary conditions. For this exploration, consider encrypting the message \( M = 45678 \) using a second-order non-homogeneous differential equation. The added complexity of a second-order equation models realistic challenges in encryption while highlighting the role of calculus.

\subsubsection*{Step 1: Define the Encryption Differential Equation}

We select the differential equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = f(x),
\]
where \( f(x) = 45678x^2 + 34125 \) encodes the message. The corresponding initial conditions are chosen as \( y(0) = 0 \) and \( y'(0) = M \), embedding \( M = 45678 \) as part of the solution.

\subsubsection*{Step 2: Solve the Differential Equation}

The general solution to this second-order equation is composed of a homogeneous part and a particular part:
\[
y(x) = y_h(x) + y_p(x),
\]
where \( y_h(x) \) solves the homogeneous equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = 0,
\]
and \( y_p(x) \) is a particular solution to the full equation.

\paragraph*{Homogeneous Solution}
For the homogeneous equation:
\[
r^2 - 5r + 6 = 0,
\]
solve the characteristic equation for roots:
\[
r = 2, \quad r = 3.
\]
Thus, the homogeneous solution is:
\[
y_h(x) = C_1e^{2x} + C_2e^{3x},
\]
where \( C_1 \) and \( C_2 \) are constants determined by the initial conditions.

\paragraph*{Particular Solution}
Assume a particular solution of the form \( y_p(x) = Ax^3 + Bx^2 + Cx + D \). Substituting \( y_p(x) \) and its derivatives into the original equation:
\[
\frac{d^2}{dx^2}(Ax^3 + Bx^2 + Cx + D) - 5\frac{d}{dx}(Ax^3 + Bx^2 + Cx + D) + 6(Ax^3 + Bx^2 + Cx + D) = 45678x^2 + 34125.
\]
Expanding and equating coefficients of \( x^n \), solve for \( A, B, C, \) and \( D \). This yields:
\[
y_p(x) = 15226x^3 - 30452x^2 + 45678x + 34125.
\]

Thus, the general solution is:
\[
y(x) = C_1e^{2x} + C_2e^{3x} + 15226x^3 - 30452x^2 + 45678x + 34125.
\]

\subsubsection*{Step 3: Apply Initial Conditions}

@@

@@

>>>>

\setlength{\parindent}{0.5in} % Sets the paragraph indentation to 0.5 inches

@@

% Actual exploration
\section*{Exploration}

\subsubsection*{Mathematical Complexity and Security}
When comparing calculus-based cryptography to the discrete logarithm problem (DLP), the differences in mathematical foundations, computational efficiency, and security are evident. Calculus-based cryptographic methods rely on solving differential equations using techniques such as substitution, partial fractions, and numerical solutions, which require an advanced understanding of calculus concepts \cite{Stewart2016}. On the other hand, DLP relies on modular arithmetic and group theory, with roots in simpler operations like exponentiation and congruences, making its mathematical basis more accessible \cite{Trappe2006}.

While calculus-based methods present an opportunity to explore sophisticated analytical tools, they are computationally intensive. This limits their efficiency in large-scale data processing compared to the well-established modular exponentiation algorithms used in DLP-based cryptographic systems, which are optimized for performance \cite{Koblitz1994}. From a security perspective, the DLP has been extensively analyzed and is considered a benchmark for cryptographic strength, as its security relies on the computational difficulty of solving discrete logarithms \cite{Trappe2006}. Conversely, calculus-based approaches, which depend on the difficulty of reverse-engineering differential equations, are less validated and lack the same level of cryptanalytic scrutiny.

% Explaining Calculus Cryptography and my algorithmic methodology to solve it %
\subsubsection*{Foundation of Calculus Cryptography}

The method of Calculus cryptography that I will be using in my comparison utilizes  differential equations as encryption mechanisms. A simple example involves encoding information as the solution to a specific differential equation. The sender encodes a message \( M \) by creating a differential equation where \( M \) is embedded within the boundary or initial conditions. For example:

@@

In this instance, \( M \) is the the initial condition. The encrypted data is transmitted in the form of \( f(x, y) \), and only someone who knows what \( M \) is can actually correctly solve the equation and recover the message. To actually use calculus cryptography, there are three key components:

\begin{enumerate}
    \item \textbf{Encryption:} Represent the message \( M \) as an initial value in a differential equation.
    \item \textbf{Transmission:} Share the function \( f(x, y) \) and any necessary boundary constraints, keeping \( M \) private.
    \item \textbf{Decryption:} Solve the differential equation using the initial condition.
\end{enumerate}

To showcase the application of calculus cryptography, I will encrypt a message \( M = 5 \) using a first-order differential equation and demonstrate the decryption process. In this example, Let \( f(x, y) = y + x^2 \), and the initial condition \( y(0) = 5 \). The differential equation is:

@@

% Explaining the Discrete Logarithmic Problem and my algorithmic methodology to solve it %
\subsubsection*{Foundation of Discrete Logarithmic Problem (DLP)}

For my comparison, I have included the Discrete Logarithmic Problem due to the heavy reliance of this problem in the current cryptographic space. As explained in the introduction, the DLP is a fundamental problem in encoding message. It involves finding an integer \( x \) such that:
\[
g^x \equiv h \pmod{p}
\]
where \( g \) is a generator, \( h \) is the resultant, and \( p \) is a prime modulus. DLP-based systems, such as Diffie-Hellman and ElGamal, derive their security from the computational difficulty of solving this congruence \cite{HandbookAppliedCryptography}. The DLP process that I will be using can be broken down into three key steps:

\begin{enumerate}
    \item \textbf{Parameter Generation:} Select \( g \) and \( p \) such that \( g \) generates a large cyclic group under modulo \( p \).
    \item \textbf{Public Sharing:} Share \( g \), \( p \), and \( g^x \) while keeping \( x \) private.
    \item \textbf{Encryption and Decryption:} Use \( x \) as the secret key for modular exponentiation operations.
\end{enumerate}

To showcase the application of the discrete logarithmic problem (DLP), I will encrypt a message \( M = 9 \) using modular exponentiation and demonstrate the decryption process. In this example, let the generator \( g = 3 \), the prime modulus \( p = 17 \), and the private key \( x = 4 \). The encryption process computes:

\[
C = g^x \pmod{p}
\]
\centering Substituting the values, I can calculate:
\[
C = 3^4 \pmod{17} = 81 \pmod{17} = 13
\]

As a result, the encrypted message, or ciphertext, is \( C = 13 \).

\vspace {0.1in}

To decrypt, the receiver needs to recover \( x \) such that:
\[
3^x \equiv 13 \pmod{17}
\]
This requires solving the DLP. Testing successive powers of \( 3 \) modulo \( 17 \), I could find that:
\[
3^4 \equiv 13 \pmod{17}
\]

\raggedright
\setlength{\parindent}{0.5in} % Sets the paragraph indentation to 0.5 inches

With this, the final result shows \( x = 4 \). Using \( x \) and the shared parameters \( g \) and \( p \), the receiver decrypts the original message \( M = 9 \). If a receiver does not have the private key \( x \), the decryption process becomes computationally infeasible, ensuring the security of the message. However, having access to \( x \) makes the decryption process computationally straightforward, successfully retrieving \( M \).

@@

\subsection*{Practical Case of Calculus Cryptography}

For this exploration, I will be encrypting the message \( M = 45678 \) using a second-order non-homogeneous differential equation. The added complexity of a second-order equation models realistic challenges in encryption while showing the role of calculus.

To begin with the 1st key component mentioned in the \textit{Foundation of Calculus Cryptography} section, I will encrypt the message. I will be selecting the differential equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = f(x),
\]
where \( f(x) = 45678x^2 + 34125 \) encodes the message. The corresponding initial conditions are chosen as \( y(0) = 0 \) and \( y'(0) = M \), embedding \( M = 45678 \) as part of the solution.

% General Solution

Next, to encrypt the message I solved the differential equation mentioned above. The \textbf{general solution} to this second-order equation is composed of a homogeneous part and a particular part:
\[
y(x) = y_h(x) + y_p(x),
\]
\centering where \( y_h(x) \) solves the homogeneous equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = 0,
\]
and \( y_p(x) \) is a particular solution to the full equation.

\vspace {0.1in}

For the homogeneous equation:
\[
r^2 - 5r + 6 = 0,
\]
I solved the characteristic equation for roots as
\[
r = 2, \quad r = 3.
\]
As a result, the homogeneous solution is:
\[
y_h(x) = C_1e^{2x} + C_2e^{3x},
\]
in which \( C_1 \) and \( C_2 \) are constants that are determined by the initial conditions set above.

\raggedright
\setlength{\parindent}{0.5in} % Sets the paragraph indentation to 0.5 inches

\vspace {0.1in}

% Particular Solution

For the \textbf{particular solution}, I used the assumption that the form was \( y_p(x) = Ax^3 + Bx^2 + Cx + D \). By substituting \( y_p(x) \) and its derivatives into the original equation, I got
\[
\frac{d^2}{dx^2}(Ax^3 + Bx^2 + Cx + D) - 5\frac{d}{dx}(Ax^3 + Bx^2 + Cx + D) + 6(Ax^3 + Bx^2 + Cx + D) = 45678x^2 + 34125.
\]
\centering Expanding and equating coefficients of \( x^n \), I solved for \( A, B, C, \) and \( D \). The result was
\[
y_p(x) = 15226x^3 - 30452x^2 + 45678x + 34125.
\]

The general solution of this is:
\[
y(x) = C_1e^{2x} + C_2e^{3x} + 15226x^3 - 30452x^2 + 45678x + 34125.
\]

\raggedright
\setlength{\parindent}{0.5in} % Sets the paragraph indentation to 0.5 inches

Next, I applied the initial conditions to determine the constants \( C_1 \) and \( C_2 \) and find the particular solution of the encrypted \( M \).

% Exploration of DLP
\subsection*{Practical Case of Discrete Logarithmic Problem (DLP)}

@@

\subsubsection*{Encryption}

@@

@@

\bibitem{HandbookAppliedCryptography}
Menezes, Alfred J., et al. \textit{Handbook of Applied Cryptography}. CRC Press, 1996.

@@

\bibitem{Stewart2016}
Stewart, James. \textit{Calculus: Early Transcendentals}. Cengage Learning, 2016.

\bibitem{Trappe2006}
Trappe, Wade, and Lawrence C. Washington. \textit{Introduction to Cryptography with Coding Theory}. Pearson, 2006.
